---
title: "Data_Tidying"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, warning = F, message = F}
library(dplyr)
library(tidyr)
catch_df<-read.csv(file='https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.302.1',
                    stringsAsFactors = FALSE)
                
head(catch_df)


```

Calling a function from a specific library: 'library_name::function_name( ... )'

cmd + shift + m %>% gives pipe operator

'gather' will create long format.

# Data cleaning
```{r}
catch_long <- catch_df %>% 
  select(-All, -notesRegCode) %>% 
  gather(key = "species", value = "catch", -Year, -Region)

head(catch_long)
```

Corrected ascii value in catch, line 401.

```{r}
catch_cleaned <- catch_long %>% 
  rename(catch_thousands = catch) %>% 
  mutate(catch_thousands = ifelse(catch_thousands == 'I', 1, catch_thousands)) %>% 
  mutate(catch_thousands = as.integer(catch_thousands)) %>% 
  mutate(catch=catch_thousands*1000)

head(catch_cleaned)
```

```{r, echo=F}

# used this code to test one value

test_catch <- as.integer(catch_cleaned$catch_thousands)

i<- which(is.na(test_catch) == T)
catch_cleaned[i, ]
```

We've just created a clean dataframe (catch_long). Removed marginal sums and notes.
Moved from wide to long (tall) format. Negative values in 'select' function remove columns.

# Split-Apply-Combine

Calculate total catch by region

```{r}
catch_total <- catch_cleaned %>% 
  group_by(Year, Region, species) %>% 
  summarize(catch_total = sum(catch))

# catch_total

catch_matt <- catch_cleaned %>% 
  group_by(species, Year) %>% 
  summarise(catch_mean = mean(catch), catch_sd=sd(catch), catch_n=n())

head(catch_matt)
```

Filter by species and something else

```{r}
catch_chinook <- catch_cleaned %>% 
  filter(species == "Chinook" & Region == "SSE" & Year > 1990) %>% 
  arrange(Year)

catch_chinook
```


# Joins

```{r}
region_defs <- read.csv(url("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/df35b.303.1", method = "libcurl"), stringsAsFactors = FALSE)

head(region_defs)
```

```{r}
region_clean <- region_defs %>% 
  select(code, mgmtArea)

head(region_clean)
```

```{r}
catch_joined <- left_join(catch_cleaned, region_clean, 
                          by = c("Region" = "code"))

head(catch_joined)
```

# Spread

Make a wide dataframe

```{r}
catch_wide <- catch_cleaned %>% 
  filter(Year > 1990) %>% 
  select(-catch_thousands) %>% 
  spread(key = Year, value = catch)

head(catch_wide)
```


# Separate function, Unite

```{r}
dates_df <- data.frame(date = c("5/24/1930",
                                "5/25/1930",
                                "5/26/1930",
                                "5/27/1930",
                                "5/28/1930"),
                       stringsAsFactors = FALSE)

dates_df # %>% 
#  separate(date, c("month", "day", "year"), "/")
```

```{r}
dates_sep <- dates_df %>% 
  separate(col = date, into = c("month","day","year"), "/")

dates_sep
```

```{r}
date_unite <- dates_sep %>% 
  unite(date_iso, year, month, day, sep = "-")

head(date_unite)
```