---
title: "Anything goes"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F)
library(dplyr)
library(tidyr)
library(ggplot2)
library(DT)
```

# Data tidying

Read in data from KNB data repository, ADFG

Alaska Department of Fish and Game, Division of Commercial Fisheries. 2017. Daily salmon escapement counts from the OceanAK database, Alaska, 1921-2017. Knowledge Network for Biocomplexity. doi:10.5063/F1S46Q6M


```{r, echo = F, results = F}
esc<-read.csv("https://knb.ecoinformatics.org/knb/d1/mn/v2/object/urn%3Auuid%3Af119a05b-bbe7-4aea-93c6-85434dcb1c5e", stringsAsFactors = F)

# head(esc)
```


# Calculating annual

separate date into yr-mon-day

calculate regional totals 

We will use 'group_by' and 'summarize' and  sum over unique regions for individual species

Also, filter for salmon species

```{r, echo = F}
annual_esc <- esc %>% 
  separate(sampleDate, into = c("year", "month", "day"), sep = "-") %>% 
  mutate(year = as.numeric(year)) %>% 
  group_by(SASAP.Region, Species,year) %>% 
  summarise(escapement = sum(DailyCount)) %>% 
  filter(Species %in% c("Chinook", "Sockeye", "Pink", "Chum", "Coho"))

write.csv(annual_esc, "annual escapement.csv", row.names = F)
  
# head(annual_esc)
```

```{r, echo = F}
datatable(annual_esc)
```
# Static plots

Use ggplot cheatsheets to look for options

```{r}
ggplot(data = annual_esc, mapping = aes(x = Species, y = escapement, fill = SASAP.Region)) + 
  geom_col()
```

Timeseries in Kodiak by species

```{r}
my_theme = theme_bw() +
  theme(plot.title = element_text(hjust = 0.5))

kodiak_esc = filter(annual_esc, SASAP.Region == "Kodiak")
ggplot(data = kodiak_esc, mapping = aes(x = year, y = escapement, color = Species)) + 
  geom_line() + 
  geom_point() +
  ylab(bquote('Escapement (number of fish)  '  ~D^15~'')) +
  xlab("Year") +
  ggtitle("Кодиак Salmon Escapement") +
  my_theme
```

you can define my_theme to remember your favorite option. Next time you plot anything, just call my_theme.

Plot timeseries of escapement by species for all regions, separately.

```{r, fig.height = 8, fig.width=6}
p1 <- ggplot(annual_esc, aes(x = year, y = escapement, color = Species)) +
  geom_line() +
  geom_point() +
  facet_wrap(~SASAP.Region, scales = "free_y", ncol = 2) +
  my_theme

p1

ggsave("region_escapement.png", plot = p1, height = 10, width = 8, units = "in")
```
# Interactive plots
